Habitat edges affect tree diversity more than biomass regeneration in a wet neotropical secondary forest

### 2022-07-11 article/merge.tex
first poster draft

\title{Habitat edges affect tree diversity more than biomass regeneration in a wet neotropical secondary forest}
\author{Nicholas Medina}
\date{scriptsize~Ecology and Evolutionary Biology, University of Michigan, 1105 N University Ave, Ann Arbor, MI, USA 48109}

@@

\hypertarget{draft-attic}{%
\section*{draft attic}\label{draft-attic}}
\addcontentsline{toc}{section}{draft attic}

This study supports previous syntheses suggesting that forest biomass and taxonomic richness regenerate more quickly / are more resilient than diversity and community composition \emph{(\protect\hyperlink{ref-poorter21}{Poorter et al., 2021}?, other?)}.
This may be due to priority effects on successional trajectories acting on/at the stem/population level \emph{{[}{]}}, variable turnover and reproduction rates among tree taxa \emph{{[}{]}}, and/or differential dispersal and establishment success based on local {[}dispersive{]} animal community \emph{{[}{]}}. \ldots{}
- mention more amundant genera patterns w/ \emph{Vochysia} and \emph{Ficus} as a key fruit-bearing taxon \emph{(\protect\hyperlink{ref-tropicalbatux2fbird}{\textbf{tropicalbat/bird?}})}?

There is increasing interest in functional trait ecology of tropical trees \emph{(\protect\hyperlink{ref-umauxf1a}{\textbf{umaña?}}\#\#)}.
This focuses on a coarse functional association with successional stage, \ldots{}
but also includes patterns with wood density {[}underlying biomass estmations{]}\ldots{}
- fig3 -- edge effect on succession generalists, but not specialists
The persistence of early-successional tree communities has been found to lower the standing biomass, likely due to lower wood density \emph{(\protect\hyperlink{ref-cuxe9sar21}{\textbf{césar21?}})}, and could therefore require a switch in focal timber species \emph{(\protect\hyperlink{ref-karsten14}{Karsten et al., 2014})}.

\hypertarget{attic}{%
\section*{attic}\label{attic}}
\addcontentsline{toc}{section}{attic}

In contrast, under intensive silviculture that prioritizes the growth of select tree crop species, forest stands can show high wood production (Ferez et al 2015) while also usually showing negative trade-offs with other services, like that with biodiversity given reliance on monoculture planting designs.

In some cases where management decisions are made primarily based on regional economic but not local ecological information, due to limited access to funded scientific reporting (Cook et al 2012), forest stands can perform very poorly all-around.

After this more extreme type of management, it remains uncertain how natural regeneration in secondary forests will proceed, how local ecosystem services will be restored, and how restoration could be managed locally (Laurance et al 2012), e.g.~based on surrounding forest proximity.

@@

Thanks to Professor Dan Perlman for initial networking; concurrent Osa Conservation managers Max Villalobos and Andrea Johnson for project guidance and supervision; Lehigh University Professor Don Morris and interns Emma Capetz and Erin Lau along with field station staff Agustín Mendoza for direct field assistance; concurrent Osa Conservation general manager Manuel Ramirez and Piro Biological Station staff Juan Carlos Cruz Díaz, Annia Barrantes, Larry Villalobos, Hansel Vargas, and visitors for field support; and peer James Fifer for discussion of early draft.

@@

\leavevmode\vadjust pre{\hypertarget{ref-karsten14}{}}%
Karsten, R.J., Meilby, H., Larsen, J.B., 2014. Regeneration and management of lesser known timber species in the {Peruvian Amazon} following disturbance by logging. Forest Ecology and Management 327, 76--85. \url{https://doi.org/10.1016/j.foreco.2014.04.035}

>>>>

\title{Habitat edges affect tree diversity more than biomass regeneration in a wet neotropical secondary forest}
\author{Nicholas Medina}
\date{Ecology and Evolutionary Biology, University of Michigan, 1105 N University Ave, Ann Arbor, MI, USA 48109}

@@

@@

Thanks to Professor Dan Perlman for initial networking; concurrent Osa Conservation managers Max Villalobos and Andrea Johnson for project guidance and supervision; Lehigh University Professor Don Morris and interns Emma Capetz and Erin Lau along with field station staff Agustín Mendoza for direct field assistance; concurrent Osa Conservation general manager Manuel Ramirez and Piro Biological Station staff Juan Carlos Cruz Díaz, Annia Barrantes, Larry Villalobos, Hansel Vargas, and visitors for field support; peer James Fifer for discussion of early draft; and Luisa Valdez for translation assistance.

@@
